---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(OncoCast)
library(knitr)
library(dplyr)
library(ggplot2)
library(ggrepel)
```

# `OncoCast` R package

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/AxelitoMartin/OncoCast.svg?branch=development)](https://travis-ci.org/AxelitoMartin/OncoCast)
[![Codecov test coverage](https://codecov.io/gh/AxelitoMartin/OncoCast/branch/development/graph/badge.svg)](https://codecov.io/gh/AxelitoMartin/OncoCast?branch=development)
[![DOI](https://zenodo.org/badge/211392347.svg)](https://zenodo.org/badge/latestdoi/211392347)
<!-- badges: end -->


## An Ensemble Learning Approach for Outcome Prediction in Precision Oncology Setting

Ensemble learner framework for survival outcome prediction and stratification for high dimension data. Originally developed for cancer genomics with delayed entry in the risk set in mind, and thus can adjust for left-truncation. 
OncoCast enables users to easily perform one or multiple machine learning survival analyses at once and explore and visualize the resulting output.

## Installing `OncoCast`

`OncoCast` has some dependencies that will be installed if they are not found in your library. When installing from github a prompt in the console may ask if you want to install the binaries for curl v4.0 instead of 3.3. There is no need to update it for the package to work properly.

``` r
install.packages("remotes")
remotes::install_github("AxelitoMartin/OncoCast")
```
If you wish to use the development version of the package please use:

``` r
remotes::install_github("AxelitoMartin/OncoCast", ref = "development")
```

## Using `OncoCast` 

We recommend users to walk through the companion website to this package before their first use of the method. They will be guided through:

- [**Generating data in a ready for analysis format**](https://axelitomartin.github.io/OncoCast/articles/OncoCast-Data.html) where the users will be guided on how to format their data to use `OncoCast`.
- [**Performing an ensemble learning run**](https://axelitomartin.github.io/OncoCast/articles/Ensemble-Learner.html) through a guide of the different machine learning algorithms available and how to create an ensemble model.
- [**Generating comprehensible results**](https://axelitomartin.github.io/OncoCast/articles/Risk-Prediction.html) with a tutorial on how to explore the predicted risk score of patients in the dataset, the prognostic power of the ensemble model, the importance of each feature of interest and how to optimize creation of risk groups in clinically meaningful subsets.
- [**Using web-based interactive applications**](https://axelitomartin.github.io/OncoCast/articles/Online-Tools.html) to simplify the exploration of results and sharing them collaborators.


## `OncoCast` Online

There exist a version of OncoCast completely web-based requiring no coding skills and minimal inputs to create and explore an ensemble model. It can be found through this [**online RShiny application**](https://axelitomartin.shinyapps.io/OncoCast/?_ga=2.133036616.733046146.1606492754-1896886074.1572897510). The user will only be asked to input the dataset they wish to study and the method they want to use to create the emsemble model.

## Introductory example

Here we show a simple example of OncoCast's capacities utilizing an internal dataset. First constructing the ensemble model using for simplicity the LASSO method:

```{r}
onco_run <- OncoCast(data = survData, family = "cox", formula = Surv(time,status)~., method = c("LASSO"),
                     runs = 50, cores = 1, save = FALSE
)
```

The output of the [`OncoCast()`](https://axelitomartin.github.io/OncoCast/reference/OncoCast.html) function can used directly in the [`getResults_OC()`](https://axelitomartin.github.io/OncoCast/reference/getResults_OC.html) function to create a set of comprehensible results to be explored.

```{r}
results <- getResults_OC(OC_object = onco_run$LASSO, data = survData)
```

This will generate the average predicted risk score for each patient in the dataset, while automatically creating risk groups through Kmeans clustering using a penalty:

```{r}
kable(results$RiskScoreSummary)
results$RiskHistogram
```

In order to assess the prognostic performance of the ensemble model we record the concordance probability index found at each iteration:

```{r}
kable(results$CPE)
```

Along with the average predicted risk score of each patients we can use as a predictor to refit a Cox proportional hazard model:

```{r}
kable(results$RiskRefitTable)
```

Another metric of interest is the variable importance that we report using a variety of interactive plots, for the penalized regression methods such as LASSO (used here) we report this variable importance in the shape of a volcano plot. We show here a static version of it due to issue with README files: 

```{r}
mod_data <- results$resultsAll %>%
  mutate(Feature = rownames(results$resultsAll))

mod_data %>%
  ggplot(aes(x = exp(MeanCoefficient), y = SelectionFrequency, label = Feature)) +
  geom_point(aes(color = .data$SelectionFrequency > 0.8), show.legend = FALSE) +
  geom_text_repel(data = mod_data %>%
              filter(SelectionFrequency > 0.8) ) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  geom_hline(yintercept = 0.8, linetype = "dashed", color = "blue") +
  theme(legend.position = "none") +
  xlab("Mean hazard ratio") + ylab("Selection Frequency") +
  theme_bw()
```

Please refer to the [**Risk_Prediction**](https://axelitomartin.github.io/OncoCast/articles/Risk-Prediction.html) tab on the package's website to see the interactive version.

We also create heatmaps sorted by predicted risk (left is low risk, right is high risk) to explore the distribution of features as a function of risk score:

```{r}
results$heatmap.sorted.bin
```

Finally we show the resulting stratification from the cuts performed above through a Kaplan-Meier plot along with a summary table to ease interpretation:

```{r, fig.height=7}
results$KM
kable(results$survivalTable)
```


We also include tools to perform validation and individual patient survival predictions, please refer to the [**Risk_Prediction**](https://axelitomartin.github.io/OncoCast/articles/Risk-Prediction.html) tab on the package's website for more information on those topics.
